openarchaeoCollaboration

### 2023-01-31 analysis/paper.qmd
Add analysis of betweenness vs. activity

```{r dependencies, include=FALSE}
library("tidyverse")
library("lubridate")
library("openarchaeoCollaboration")
library("khroma")
library("gt")
library("fs")
library("here")
library("tidygraph")
library("ggraph")
library("ggiraph")
```

@@

@@

```{r plot-gh_commits-cum}
oarch %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  mutate(date = floor_date(datetime, "week")) %>%
  mutate(lumped_category = recode(category,
                                  "Specifications etc." = "Documents",
                                  "Products" = "Documents",
                                  "Guides" = "Documents",
                                  "Lists and datasets" = "Documents",
                                  .default = category)) %>%
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cat_total = sum(n_commits)) %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = fct_reorder(lumped_category, cat_total),
             colour = fct_reorder(lumped_category, cat_total))) +
  geom_area() +
  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
                              ymd("2021-05-31", tz = "UTC")),
                   date_breaks = "2 years",
                   date_labels = "%Y") +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_bright(guide = NULL) +
  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Growth of open archaeology",
    subtitle = "Measured by cumulative commits in GitHub repositories",
    caption = "Data: open-archaeo.info and GitHub API",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_oarch()

@@

```{r summary-graph-user-user}
oarch_graph_user_user |>
  mutate(degree = centrality_degree(), 
         betweenness = centrality_betweenness(weights = n), 
         strength = centrality_degree(weights = n)) |>
  filter(degree > 0) |> 
  data.frame() |> 
  View()
```

```{r plot-graph-user-user}
oarch_graph_user_user |>
  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
  filter(degree > 0) |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  # geom_node_point(aes(colour = category)) +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common repositories")
```

>>>>

```{r dependencies, include=FALSE}
library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("gt")
library("here")
library("khroma")
library("lubridate")
library("openarchaeoCollaboration")
library("patchwork")
library("tidygraph")
library("tidyverse")
```

@@

### Who's doing what?

```{r user-contributors}
oarch |>
  unnest(gh_contribs) |>
  drop_na(contributor) |>
  group_by(contributor) |>
  mutate(
    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
  ) |>
  group_by(contributor, repo_owner) |>
  summarise(repos = n(), contribs = sum(contributions)) |>
  pivot_wider(c(contributor), names_from = repo_owner, 
              values_from = c(repos, contribs), values_fill = 0) |>
  mutate(
    repos = repos_other + repos_own,
    contribs = contribs_other + contribs_own,
    .after = contributor
  ) ->
  oarch_contributors
```

@@

```{r plot-gh-cum}
oarch %>% 
  mutate(
    lumped_category = recode(category,
                             "Specifications etc." = "Documents",
                             "Products" = "Documents",
                             "Guides" = "Documents",
                             "Lists and datasets" = "Documents",
                             .default = category),
    lumped_category = factor(lumped_category, c("Scripts",
                                                "Documents",
                                                "Packages and libraries",
                                                "Standalone software")) ) ->
  oarch_lumped

p_gh_commits_cum <- oarch_lumped %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  mutate(date = floor_date(datetime, "week")) %>%
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = lumped_category,
             colour = lumped_category,
             label = lumped_category)) +
  geom_area() +
  annotate("segment", 
           x = ymd("2008-04-01", tz = "UTC"), 
           xend = ymd("2008-04-01", tz = "UTC"),
           y = 16000, yend = 4000,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", 
           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
           label = "GitHub\nlaunched",
           vjust = 0) +
  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_bright(guide = guide_none()) +
  scale_fill_bright() +
  labs(
    y = "cumulative commits",
    x = NULL, fill = NULL
  ) +
  theme_oarch() +
  theme(legend.position = "top")

# TODO: extend to end
p_gh_repos_cum <- oarch_lumped %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  group_by(repo) %>% 
  filter(datetime == min(datetime)) %>% # Only first commits
  mutate(date = floor_date(datetime, "week")) %>%
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cat_total = sum(n_commits)) %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = lumped_category,
             colour = lumped_category)) +
  geom_area() +
  annotate("segment",
           x = ymd("2008-04-01", tz = "UTC"),
           xend = ymd("2008-04-01", tz = "UTC"),
           y = 40, yend = 10,
           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text",
           x = ymd("2008-04-01", tz = "UTC"), y = 50,
           label = "GitHub\nlaunched",
           vjust = 0) +
  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_bright(guide = guide_none()) +
  scale_fill_bright(guide = guide_none()) +
  labs(
    y = "cumulative repositories",
    x = NULL, fill = NULL
  ) +
  theme_oarch()

p_gh_commits_cum / p_gh_repos_cum

@@

```{r summary-graph-user-user}
oarch_graph_user_user |>
  mutate(degree = centrality_degree(), 
         betweenness = centrality_betweenness(weights = n), 
         strength = centrality_degree(weights = n)) |>
  filter(degree > 0) |> 
  data.frame()
```

```{r plot-graph-user-user}
oarch_graph_user_user |>
  mutate(betweenness = centrality_betweenness(weights = n)) ->
  oarch_graph_user_user

oarch_graph_user_user |> 
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  # geom_node_point() +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common repositories")
```

```{r plot-user-betweenness-dist}
oarch_graph_user_user |>
  mutate(
    betweenness = centrality_betweenness(weights = n)
  ) |>
  data.frame() |>
  ggplot(aes(x = betweenness)) +
  geom_density(bw = "nrd") +
  scale_x_log10(labels = scales::number)
```

```{r plot-user-betweenness-activity}
oarch_graph_user_user |>
  mutate(
    degree = centrality_degree(), 
    betweenness = centrality_betweenness(weights = n), 
    strength = centrality_degree(weights = n)
  ) |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm")
```
